---
title: "demo"
author: "Ugne Milasiunaite"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
library(icenReg)
library(flexsurv)
```

# Heper functions

## Generating the data set

```{r}
#' Generate time-to-event data from the Weibull distribution
#' 
#' @param n A named list providing the register sizes
#' @param proportions A named list of named lists of the proportions of people receiving each treatment for each register
#' @param pars A named list of Weibull dist. parameter values for each treatment
#' @return A data frame with four columns: survival time, probability, treatment type, register
generate_weibull <- function(n, proportions, pars) {
  # Save names for later use
  regs <- names(proportions)
  treatments <- names(pars)

  # Calculation of sizes
  k <- sapply(regs, function(i)
    sapply(treatments, function(tr)
      round(n[[i]] * proportions[[i]][[tr]])))
  
  # Generate time-to-event data
  t.vec <- sapply(treatments, function(i) 
    rweibull(sum(k[i,]), shape = pars[[i]][[1]], scale = pars[[i]][[2]]))
  
  # Create the combined data set
  df <- do.call(rbind, lapply(treatments, function(tr)
    data.frame(
      "time" = t.vec[,tr],
      "survival" = 1 - pweibull(t.vec[,tr], pars[[tr]][[1]], pars[[tr]][[2]]),
      "treatment" = as.factor(tr),
      "register" = as.factor(rep(regs, k[tr, ]))
    )))
  
  return(df)
}


#' Generate status, entry and exit dates given survival time in days
#' 
#' @param days Vector of survival time in days
#' @param duration Duration of enrollment in days
#' @param followup Duration of follow-up from end of enrollment in days
#' @param prob Probability of being lost-to-follow-up for each individual
#' @return A data frame with
generate_censoring <- function(days, duration, followup, prob) {
  # Generate entry times
  K <- length(days)
  lost <- rbinom(K, 1, prob)
  t0 <- if(duration > 0) ceiling(runif(K, 0, duration)) else 0
  
  # Create the data frame
  df <- data.frame(
    "status" = as.factor(ifelse(
      lost, "Lost", ifelse(days > t0 + followup, "Alive", "Dead")
    )))
  df["exit"] <- ifelse(
    df$status == "Alive",
    t0 + followup,
    ifelse(!lost, days, ceiling(runif(K, 0, pmin(t0 + followup, days))))
  )
  
  return(df)
}


#' Generate lower and upper limits for interval-censored data
#' 
#' @param exit Date of exit
#' @param exact Binary indicator whether the date should be left exact
#' @param infinite Binary indicator if the upper limit is infinite
#' @return
generate_intervals <- function(exit, exact, infinite) {
  L <- ifelse(exact | infinite, exit, exit - floor(runif(length(exact), 0, pmin(exit, 30))))
  U <- ifelse(exact, exit, ifelse(infinite, Inf, L + 30))
  return(data.frame("lower" = L, "upper" = U))
}


#' Generate the data set
#' 
#' @param n A named list providing the register sizes
#' @param proportions A named list of named lists of the proportions of people receiving each treatment for each register
#' @param pars A named list of Weibull dist. parameter values for each treatment
#' @param duration Duration of enrollment in days
#' @param followup Duration of follow-up from end of enrollment in days
#' @param prob Probability of being lost-to-follow-up in each register
generate_df <- function(n, proportions, pars, duration, followup, prob) {
  df <- generate_weibull(n, proportions, pars)
  df.censoring <- generate_censoring(ceiling(df$time), duration, followup, prob[df$register])
  df.intervals <- generate_intervals(
    df.censoring$exit,
    exact = df$register == "1" & df.censoring$status == "Dead",
    infinite = df.censoring$status != "Dead"
  )
  return(do.call(cbind, list(df, df.censoring, df.intervals)))
}

```

# Simulation

```{r}
set.seed(311)

N <- 4000
N.prop <- 1/2
N1 <- N*N.prop

# Description of the setting
sizes <- c("1" = N1, "2" = N - N1)
treat.prop <- list("1" = list("A" = 0.5, "B" = 0.5), 
                   "2" = list("A" = 0.5, "B" = 0.5))
parameters <- list("A" = c(0.5, 2000), "B" = c(0.5, 2500))
lost.prop <- c("1" = 0.01, "2" = 0.01)
duration <- 365
followup <- 1826

tr.names <- names(parameters)
cols <- c("A" = "#009292", "B" = "#B66DFF")

# Create the observed data set
df.obs <- generate_df(sizes, treat.prop, parameters, duration, followup, lost.prop)
df.obs["obs"] <- ifelse(
  df.obs$register == 1,
  df.obs$exit,
  ifelse(df.obs$status == "Dead", df.obs$upper, df.obs$exit)
)

df.obs[c((N1/2-20):(N1/2-19), (N1-19):(N1-20), (N1+19):(N1+20), (N-20):(N-19)),]
```

Checks for data

```{r}
# exit, lower, obs cannot be more than duration + followup
all(c(df.obs$exit, df.obs$lower, df.obs$obs) <= duration + followup)
# lower cannot be bigger than upper
all(df.obs$lower <= df.obs$upper)
# people per treatment
(N.tr <- table(df.obs$treatment))
# people per treatment per register
table(df.obs[c("register", "treatment")])
# visualize the distribution of times
dtimes <- table(list("time.int" = cut(df.obs$time, c(0, 1, 5, 30, 60, 180, 365, 1095, 1826, Inf)), "treatment" = df.obs$treatment))
cbind(dtimes, "propA" = dtimes[, "A"] / N.tr["A"], "propB" = dtimes[, "B"] / N.tr["B"])
# visualize status
matrix(table(df.obs[c("status", "treatment", "register")]), nrow = 3, ncol = 4,
       dimnames = list(c("Alive", "Dead", "Lost"), c("r1A", "r1B", "r2A", "r2B")))
```

# Model fitting

```{r}
# models for right-censored data
fit_imputed <- function(Surv.obj, df, tr.cols = cols) {
  tr.names <- names(tr.cols)
  
  # 1. Kaplan-Meier estimator
  fit.KM <- survfit(eval(Surv.obj) ~ treatment, data = df, conf.type = "log-log")
  plot(fit.KM, main = "Kaplan-Meier survival curves", col = tr.cols,
     xlab = "Days", ylab = "Survival")
  points(df$time, df$survival, cex = 0.1, col = "#99999966")
  legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
  
  # Check for proportional hazards
  plot(fit.KM, fun = "cloglog", col = cols, xlab = "Days", ylab = "log(-log(Survival))")
  legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
  
  # 2. Cox proportional hazards model
  fit.cox <- coxph(eval(Surv.obj) ~ treatment, data = df)
  
  # Check for proportional hazards
  res <- cox.zph(fit.cox, terms = TRUE)
  print(res)
  plot(res, resid = TRUE, main = "Schoenfeld residual plot")
  
  # 3. Weibull AFT/PH model
  fit.aft <- survreg(eval(Surv.obj) ~ treatment, data = df, dist = "weibull")
  
  # 4. Weibull AFT/PH stratified model
  fit.aft.strat <- survreg(eval(Surv.obj) ~ strata(treatment), data = df, dist = "weibull")
  
  return(list(fit.KM, fit.cox, fit.aft, fit.aft.strat))
}
```

## If everyone's times were accurate to the day

```{r}
models.day <- fit_imputed(expression(Surv(exit, status == "Dead")), df.obs)
```

Predict survival probabilities with 95% confidence intervals

```{r}
predict_surv_imputed <- function(D, models) {
  KM <- models[[1]]$surv[models[[1]]$time == D]
  KM.lower <- models[[1]]$lower[models[[1]]$time == D]
  KM.upper <- models[[1]]$upper[models[[1]]$time == D]
  COX <- predict(
    models[[2]],
    data.frame(exit = D, obs = D, obs.mid = D, status = "Dead", 
               treatment = c("A", "B")),
    type = "survival",
    se.fit = TRUE
  )
  COX.lower <- COX$fit - qnorm(0.975) * COX$se.fit
  COX.upper <- COX$fit + qnorm(0.975) * COX$se.fit
  AFT <- 1 - pweibull(D, 
                      shape = 1 / models[[3]]$scale,
                      scale = c(exp(models[[3]]$coefficients[[1]]),
                                exp(sum(models[[3]]$coefficients))))
  AFT.AIC <- - 2*models[[3]]$loglik[2] + 2*length(models[[3]]$coefficients)
  AFT.STRAT <- 1 - pweibull(D,
                            shape = 1 / models[[4]]$scale,
                            scale = exp(models[[4]]$coefficients))
  AFT.STRAT.AIC <- - 2*models[[4]]$loglik[2] + 2*length(models[[4]]$coefficients)
  tbl <- data.frame(
    "A" = c(KM[1], COX$fit[1], AFT[1], AFT.STRAT[1]),
    "A.lower" = c(KM.lower[1], COX.lower[1], NA, NA),
    "A.upper" = c(KM.upper[1], COX.upper[1], NA, NA),
    "B" = c(KM[2], COX$fit[2], AFT[2], AFT.STRAT[2]),
    "B.lower" = c(KM.lower[2], COX.lower[2], NA, NA),
    "B.upper" = c(KM.upper[2], COX.upper[2], NA, NA),
    "AIC" = c(NA, NA, AFT.AIC, AFT.STRAT.AIC),
    row.names = c("KM", "COX", "AFT", "AFT.STRAT")
  )
  return(tbl)
}
```

```{r}
surv.day <- predict_surv_imputed(1826, models.day)
```

## If we take the upper limit as exact time for register 2

```{r}
models.upp <- fit_imputed(expression(Surv(obs, status == "Dead")), df.obs)
```

```{r}
surv.upp <- predict_surv_imputed(1826, models.upp)
```

## If we take the upper limit - 15 (middle of the month) as exact time for register 2

```{r}
df.obs["obs.mid"] <- ifelse(df.obs$register == "2" & df.obs$status == "Dead", 
                            df.obs$obs - 15, df.obs$obs)
models.mid <- fit_imputed(expression(Surv(obs.mid, status == "Dead")), df.obs)
```

```{r}
surv.mid <- predict_surv_imputed(1826, models.mid)
```

## Interval censoring methods

```{r}
fit_interval <- function(Surv.obj, df, tr.cols = cols, ph = FALSE) {
  tr.names <- names(tr.cols)
  
  # 1. Non-Parametric Maximum Likelihood Estimator (NPMLE) aka Turnbull’s Estimator
  fit.NPMLE <- ic_np(eval(Surv.obj) ~ treatment, data = df)
  plot(
    fit.NPMLE,
    main = "NPMLE curves",
    col = tr.cols,
    xlab = "Days",
    ylab = "Survival"
  )
  points(df$time, df$survival, cex = 0.1, col = "#99999966")
  legend(
    "bottomright",
    lty = c(1, 1),
    bty = "n",
    col = tr.cols,
    legend = tr.names
  )
  
  # 2. Semi-Parametric proportional hazards model
  # only fit if the hazards are proportional since otherwise won't converge
  if(ph) {
    fit.semi.ph <- ic_sp(eval(Surv.obj) ~ treatment, data = df, model = "ph")
  } else {
    fit.semi.ph <- NA
  }
  
  # 3. Semi-Parametric proportional odds model
  fit.semi.po <- ic_sp(eval(Surv.obj) ~ treatment, data = df, model = "po")
  
  # 4. Weibull AFT/PH
  fit.aft <- survreg(eval(Surv.obj) ~ treatment, data = df, dist = "weibull")
  
  # 5. Weibull AFT/PH stratified
  fit.aft.strat <- survreg(
    eval(Surv.obj) ~ strata(treatment),
    data = df,
    dist = "weibull"
  )
  
  # 6. Weibull AFT/PH icenreg
  fit.aft.ic <- ic_par(
    eval(Surv.obj) ~ treatment,
    data = df,
    model = "aft",
    dist = "weibull",
    weights = NULL
  )
  
  return(list(fit.NPMLE, fit.semi.ph, fit.semi.po, fit.aft, fit.aft.strat, fit.aft.ic))
}
```

Fit the models

```{r}
models.int <- fit_interval(
  expression(Surv(lower, upper, type = "interval2")), 
  df.obs
)
```

```{r}
predict_surv_interval <- function(D, models) {
  survA <- models[[1]]$scurves$A
  survB <- models[[1]]$scurves$B
  NPMLE <- c(survA$S_curves[[1]][survA$Tbull_ints[, 2] >= D][1],
             survB$S_curves[[1]][survB$Tbull_ints[, 2] >= D][1])
  if(class(models[[2]]) == "ic_ph") {
    SEMI.PH <- 1 - getFitEsts(models[[2]], 
                              newdata = data.frame(treatment = c("A", "B")), 
                              q = 1826)
  } else {
    SEMI.PH <- c(NA, NA)
  }
  if(class(models[[3]]) == "ic_po"){
    SEMI.PO <- 1 - getFitEsts(models[[3]], 
                            newdata = data.frame(treatment = c("A", "B")), 
                            q = 1826)
  } else {
    SEMI.PO <- c(NA, NA)
  }
  AFT <- 1 - pweibull(D, 
                      shape = 1 / models[[4]]$scale,
                      scale = c(exp(models[[4]]$coefficients[[1]]),
                                exp(sum(models[[4]]$coefficients))))
  AFT.AIC <- - 2*models[[4]]$loglik[2] + 2*length(models[[4]]$coefficients)
  AFT.STRAT <- 1 - pweibull(D,
                            shape = 1 / models[[5]]$scale,
                            scale = exp(models[[5]]$coefficients))
  AFT.STRAT.AIC <- - 2*models[[5]]$loglik[2] + 2*length(models[[5]]$coefficients)
  tbl <- data.frame(
    "A" = c(NPMLE[1], SEMI.PH[1], SEMI.PO[1], AFT[1], AFT.STRAT[1]),
    "A.lower" = c(NA, NA, NA, NA, NA),
    "A.upper" = c(NA, NA, NA, NA, NA),
    "B" = c(NPMLE[2], SEMI.PH[2], SEMI.PO[2], AFT[2], AFT.STRAT[2]),
    "B.lower" = c(NA, NA, NA, NA, NA),
    "B.upper" = c(NA, NA, NA, NA, NA),
    "AIC" = c(NA, NA, NA, AFT.AIC, AFT.STRAT.AIC),
    row.names = c("NPMLE", "SEMI.PH", "SEMI.PO", "AFT", "AFT.STRAT")
  )
  return(tbl)
}

surv.int <- predict_surv_interval(1826, models.int)
```

## Proposed method

```{r}
#' Fit a distribution using the proposed approach
#' 
#' @param df A data frame with study data
#' @param dist The name of the parametric distribution to fit
#' @return A fitted cumulative distribution function of the modelling variable
proposed_approach <- function(df, dist) {
  fy <- function(l, u, b, x) Fz(u, b, x) - Fz(l, b, x)
  logL <- function(b, l, u, x) -sum(log(fy(l, u, b, x)))
  
  # Create coarsening intervals
  l <- ifelse(df$lower - df$upper == 0, df$lower - 1, df$lower)
  u <- df$upper
  # Covariate vector
  x <- df$treatment
  
  if (dist %in% c("exp", "exponential")) {
    Fz <- function(z, b, x) {
      (x == "A")*pexp(z, b[1]) + (x == "B")*pexp(z, b[2])
    }
    mle <- nlm(logL, c(1/100, 1/100), l, u, x, hessian = TRUE)
  } else if (dist == "gamma") {
    Fz <- function(z, b, x) {
      (x == "A")*pgamma(z, b[1], b[2]) + (x == "B")*pgamma(z, b[3], b[4])
    }
    mle <- nlm(logL, c(1, 1/100, 1, 1/100), l, u, x, hessian = TRUE)
  } else if (dist == "weibull") {
    Fz <- function(z, b, x) {
      (x == "A")*pweibull(z, exp(b[1]), exp(b[2])) + 
        (x == "B")*pweibull(z, exp(b[3]), exp(b[4]))
    }
    mle <- nlm(logL, c(-0.2, 6, -0.2, 6), l, u, x, hessian = TRUE)
  }
  stopifnot("The optimizer failed." = mle$minimum > 1)
  f <- function(z, x) Fz(z, mle$estimate, x)
  attr(f, "AIC") <- 2*mle$minimum + 2*length(mle$estimate)
  attr(f, "mle") <- mle$estimate
  return(f)
}

#' Fit a distribution using the generalised version of the proposed approach
#' 
#' @param df A data frame with study data
#' @return A fitted cumulative distribution function of the modelling variable
generalized_approach <- function(df) {
  # Define the pre-transform
  s <- function(z) qlogis(pweibull(z, 0.5, 2000))
  # Create the coarsening limits
  sl <- s(ifelse(df$lower - df$upper == 0, df$lower - 1, df$lower))
  su <- s(df$upper)
  # Define the modelling distribution and polynomial approximation
  Fz <- function(z, b, x) plogis(g(z, b, x))
  g <- function(z, b, x) {
    (b[1] + b[2]*(x == "B")) + z*(b[3] + b[4]*(x == "B"))
  }
  fy <- function(l, u, b, x) Fz(u, b, x) - Fz(l, b, x)
  logL <- function(b, l, u, x) -sum(log(fy(l, u, b, x)))
  # Estimate the parameters
  mle <- nlm(logL, rep(1, 4), u=su, l=sl, x=df$treatment)
  stopifnot("The optimizer failed." = mle$minimum > 1)
  # Save the fitted function
  f <- function(z, x) plogis(g(s(z), mle$estimate, x))
  attr(f, "AIC") <- 2*mle$minimum + 2*length(mle$estimate)
  return(f)
}

wilsonCI <- function(p.hat, n, alpha=0.05) {
  z <- qnorm(1 - alpha/2)
  lower <- (p.hat + z^2/(2*n))/(1 + z^2/n) - 
    z*sqrt(p.hat*(1 - p.hat)/n + z^2/(4*n^2))/(1 + z^2/n)
  upper <- (p.hat + z^2/(2*n))/(1 + z^2/n)+
    z*sqrt(p.hat*(1 - p.hat)/n + z^2/(4*n^2))/(1 + z^2/n)
  return(c("lower"=lower, "upper"=upper))
}
```

Fit models and get predictions

```{r}
suppressWarnings(pdist_weibull <- proposed_approach(df.obs, "weibull"))
suppressWarnings(pdist_gamma <- proposed_approach(df.obs, "gamma"))
suppressWarnings(pdist_ge <- generalized_approach(df.obs))

predict_surv_PA <- function(pdist, sizes, D) {
  p.A <- pdist(D, "A")
  p.B <- pdist(D, "B")
  CI.A <- wilsonCI(1-p.A, sizes[["A"]])
  CI.B <- wilsonCI(1-p.B, sizes[["B"]])
  AIC <- attr(pdist, "AIC")
  return(c(1-p.A, CI.A, 1-p.B, CI.B, AIC))
}

surv.pa <- matrix(
  sapply(
    list(pdist_weibull, pdist_gamma, pdist_ge),
    predict_surv_PA,
    D = 1826,
    sizes = N.tr
  ),
  nrow = 3,
  ncol = 7,
  byrow = TRUE,
  dimnames = list(
    c("weibull", "gamma", "generalised"),
    c("A", "A.lower", "A.upper", "B", "B.lower", "B.upper", "AIC")
  )
)
```

## Flexible parametric model

```{r}
fit_flexible <- function(Surv.obj, df, tr.cols = cols) {
  tr.names <- names(tr.cols)
  
  # 1. Splines
  fit.fspl0 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 0)
  fit.fspl01 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 0, anc = list(gamma1 = ~ treatment))
  fit.fspl1 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 1)
  fit.fspl11 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 1, anc = list(gamma1 = ~ treatment))
  fit.fspl12 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 1, anc = list(gamma1 = ~ treatment, gamma2 = ~ treatment))
  fit.fspl2 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 2)
  fit.fspl21 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 2, anc = list(gamma1 = ~ treatment))
  fit.fspl22 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 2, anc = list(gamma1 = ~ treatment, gamma2 = ~ treatment))
  fit.fspl23 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 2, anc = list(gamma1 = ~ treatment, gamma2 = ~ treatment, gamma3 = ~ treatment))
  fit.fspl3 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 3)
  fit.fspl31 <- flexsurvspline(eval(Surv.obj) ~ treatment, data = df, scale = "hazard", k = 3, anc = list(gamma1 = ~ treatment))
  
  # 2. Parametric
  fit.fpar <- flexsurvreg(eval(Surv.obj) ~ treatment, data = df, dist = "weibull")
  fit.fpar1 <- flexsurvreg(eval(Surv.obj) ~ treatment, data = df, dist = "weibull", anc = list(shape = ~ treatment))
  
  return(list(fit.fspl0, fit.fspl01, fit.fspl1, fit.fspl11, fit.fspl12, fit.fspl2, fit.fspl21, fit.fspl22, fit.fspl23, fit.fspl3, fit.fspl31, fit.fpar, fit.fpar1))
}

suppressWarnings(models.fpm <- fit_flexible(
  expression(Surv(lower, upper, type = "interval2")), 
  df.obs))

predict_surv_flexible <- function(D, models) {
  tbl <- lapply(
    models, 
    predict, 
    newdata = data.frame(treatment = c("A", "B")), 
    type = "survival",
    conf.int = TRUE,
    times = D
  )
  
  tbl <- cbind(t(sapply(tbl, function(x) c(x[1,2:4], x[2,2:4]))), 
               sapply(models, function(m) m$AIC))
  rownames(tbl) <- c("FPM.SPL0", "FPM.SPL01", "FPM.SPL1", "FPM.SPL11", "FPM.SPL12", "FPM.SPL2", "FPM.SPL21", "FPM.SPL22", "FPM.SPL23", "FPM.SPL3", "FPM.SPL31", "FPM.PAR", "FPM.PAR1")
  colnames(tbl) <- c("A", "A.lower", "A.upper", "B", "B.lower", "B.upper", "AIC")
  return(tbl)
}

surv.fpm <- predict_surv_flexible(1826, models.fpm)
```

# Comparisons

```{r}
true1826 <- c("A" = 1 - pweibull(1826, parameters[["A"]][[1]], parameters[["A"]][[2]]), 
              "B" = 1 - pweibull(1826, parameters[["B"]][[1]], parameters[["B"]][[2]]))
tbl <- as.matrix(rbind(c(true1826[[1]], NA, NA, true1826[[2]], NA, NA, NA), surv.day, surv.upp, surv.mid, surv.int, surv.pa, surv.fpm))
tbl
```



