---
title: "demo"
author: "Ugne Milasiunaite"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
library(icenReg)
```

# Heper functions

## Distribution and hazard functions for (stratified) Weibull distribution

```{r}
#' Stratified distribution function with no explanatory variables
#' 
#' @param v Random vector
#' @param x Treatment indicator
#' @return Vector of cumulative probabilities
pdist_strat <- function(v, x, pars) {
  lst <- lapply(names(pars), function(tr) 
    pweibull(v[x == tr], shape = pars[[tr]][[1]], scale = pars[[tr]][[2]]))
  return(do.call(c, lst))
}


#' Stratified hazard function with no explanatory variables
#' 
#' @param v Random vector
#' @param x Treatment indicator
#' @return Vector of hazards
hazard_strat <- function(v, x, pars) {
  lst <- lapply(names(pars), function(tr)
    dweibull(v[x == tr], pars[[tr]][[1]], pars[[tr]][[2]]) / (1 - pdist_strat(v[x == tr], tr, pars)))
  return(do.call(c, lst))
}
```

## Generating the data set

```{r}
#' Generate time-to-event data from the Weibull distribution
#' 
#' @param n A named list providing the register sizes
#' @param proportions A named list of named lists of the proportions of people receiving each treatment for each register
#' @param pars A named list of Weibull dist. parameter values for each treatment
#' @return A data frame with four columns: survival time, probability, treatment type, register
generate_weibull <- function(n, proportions, pars) {
  # Save names for later use
  regs <- names(proportions)
  treatments <- names(pars)

  # Calculation of sizes
  k <- sapply(regs, function(i)
    sapply(treatments, function(tr)
      round(n[[i]] * proportions[[i]][[tr]])))
  
  # Generate time-to-event data
  t.vec <- sapply(treatments, function(i) 
    rweibull(sum(k[i,]), shape = pars[[i]][[1]], scale = pars[[i]][[2]]))
  
  # Create the combined data set
  df <- do.call(rbind, lapply(treatments, function(tr)
    data.frame(
      "time" = t.vec[,tr],
      "survival" = 1 - pweibull(t.vec[,tr], pars[[tr]][[1]], pars[[tr]][[2]]),
      "treatment" = as.factor(tr),
      "register" = as.factor(rep(regs, k[tr, ]))
    )))
  
  return(df)
}


#' Generate status, entry and exit dates given survival time in days
#' 
#' @param days Vector of survival time in days
#' @param duration Duration of enrollment in days
#' @param followup Duration of follow-up from end of enrollment in days
#' @param prob Probability of being lost-to-follow-up for each individual
#' @return A data frame with
generate_censoring <- function(days, duration, followup, prob) {
  # Generate entry times
  K <- length(days)
  lost <- rbinom(K, 1, prob)
  t0 <- if(duration > 0) ceiling(runif(K, 0, duration)) else 0
  
  # Create the data frame
  df <- data.frame(
    "status" = as.factor(ifelse(
      lost, "Lost", ifelse(days > t0 + followup, "Alive", "Dead")
    )))
  df["exit"] <- ifelse(
    df$status == "Alive",
    t0 + followup,
    ifelse(!lost, days, ceiling(runif(K, 0, pmin(t0 + followup, days))))
  )
  
  return(df)
}


#' Generate lower and upper limits for interval-censored data
#' 
#' @param exit Date of exit
#' @param exact Binary indicator whether the date should be left exact
#' @param infinite Binary indicator if the upper limit is infinite
#' @return
generate_intervals <- function(exit, exact, infinite) {
  L <- ifelse(exact | infinite, exit, exit - floor(runif(length(exact), 0, pmin(exit, 30))))
  U <- ifelse(exact, exit, ifelse(infinite, Inf, L + 30))
  return(data.frame("lower" = L, "upper" = U))
}


#' Generate the data set
#' 
#' @param n A named list providing the register sizes
#' @param proportions A named list of named lists of the proportions of people receiving each treatment for each register
#' @param pars A named list of Weibull dist. parameter values for each treatment
#' @param duration Duration of enrollment in days
#' @param followup Duration of follow-up from end of enrollment in days
#' @param prob Probability of being lost-to-follow-up in each register
generate_df <- function(n, proportions, pars, duration, followup, prob) {
  df <- generate_weibull(n, proportions, pars)
  df.censoring <- generate_censoring(ceiling(df$time), duration, followup, prob[df$register])
  df.intervals <- generate_intervals(
    df.censoring$exit,
    exact = df$register == "1" & df.censoring$status == "Dead",
    infinite = df.censoring$status != "Dead"
  )
  return(do.call(cbind, list(df, df.censoring, df.intervals)))
}

```

# Simulation

```{r}
set.seed(311)

N <- 4000
N.prop <- 1/2

# Description of the setting
sizes <- c("1" = N*N.prop, "2" = N*(1-N.prop))
treat.prop <- list("1" = list("A" = 0.5, "B" = 0.5), 
                   "2" = list("A" = 0.5, "B" = 0.5))
parameters <- list("A" = c(0.5, 2000), "B" = c(0.6, 2500))
lost.prop <- c("1" = 0.01, "2" = 0.01)
duration <- 365
followup <- 1826

tr.names <- names(parameters)
cols <- c("A" = "#009292", "B" = "#B66DFF")

# Create the observed data set
df.obs <- generate_df(sizes, treat.prop, parameters, duration, followup, lost.prop)
df.obs["obs"] <- ifelse(
  df.obs$register == 1,
  df.obs$exit,
  ifelse(df.obs$status == "Dead", df.obs$upper, df.obs$exit)
)

df.obs[c(940:941, 1200:1201, 2040:2041, 3200:3201),]
```

Checks for data

```{r}
# exit, lower, obs cannot be more than duration + followup
all(c(df.obs$exit, df.obs$lower, df.obs$obs) <= duration + followup)
# lower cannot be bigger than upper
all(df.obs$lower <= df.obs$upper)
```

# Model fitting

```{r}
# models for right-censored data
fit_imputed <- function(Surv.obj, df, tr.cols = cols) {
  tr.names <- names(tr.cols)
  
  # 1. Kaplan-Meier estimator
  fit.KM <- survfit(eval(Surv.obj) ~ treatment, data = df, conf.type = "log-log")
  plot(fit.KM, main = "Kaplan-Meier survival curves", col = tr.cols,
     xlab = "Days", ylab = "Survival")
  points(df$time, df$survival, cex = 0.1, col = "#99999966")
  legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
  
  # Check for proportional hazards
  plot(fit.KM, fun = "cloglog", col = cols, xlab = "Days", ylab = "log(-log(Survival))")
  legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
  
  # 2. Cox proportional hazards model
  fit.cox <- coxph(eval(Surv.obj) ~ treatment, data = df)
  
  # Check for proportional hazards
  res <- cox.zph(fit.cox, terms = TRUE)
  print(res)
  plot(res, resid = TRUE, main = "Schoenfeld residual plot")
  
  # 3. Weibull AFT/PH model
  fit.aft <- survreg(eval(Surv.obj) ~ treatment, data = df, dist = "weibull")
  
  # 4. Weibull AFT/PH stratified model
  fit.aft.strat <- survreg(eval(Surv.obj) ~ strata(treatment), data = df, dist = "weibull")
  
  return(list(fit.KM, fit.cox, fit.aft, fit.aft.strat))
}
```

## If everyone's times were accurate to the day

```{r}
models.day <- fit_imputed(expression(Surv(exit, status == "Dead")), df.obs)
```

Predict survival probabilities with 95% confidence intervals

```{r}
predict_surv_imputed <- function(D, models) {
  KM <- models[[1]]$surv[models[[1]]$time == D]
  KM.lower <- models[[1]]$lower[models[[1]]$time == D]
  KM.upper <- models[[1]]$upper[models[[1]]$time == D]
  COX <- predict(
    models[[2]],
    data.frame(exit = D, obs = D, obs.mid = D, status = "Dead", 
               treatment = c("A", "B")),
    type = "survival",
    se.fit = TRUE
  )
  COX.lower <- COX$fit - qnorm(0.975) * COX$se.fit
  COX.upper <- COX$fit + qnorm(0.975) * COX$se.fit
  AFT <- 1 - pweibull(D, 
                      shape = 1 / models[[3]]$scale,
                      scale = c(exp(models[[3]]$coefficients[[1]]),
                                exp(sum(models[[3]]$coefficients))))
  AFT.STRAT <- 1 - pweibull(D,
                            shape = 1 / models[[4]]$scale,
                            scale = exp(models[[4]]$coefficients))
  tbl <- data.frame(
    "A" = c(KM[1], COX$fit[1], AFT[1], AFT.STRAT[1]),
    "A.lower" = c(KM.lower[1], COX.lower[1], NA, NA),
    "A.upper" = c(KM.upper[1], COX.upper[1], NA, NA),
    "B" = c(KM[2], COX$fit[2], AFT[2], AFT.STRAT[2]),
    "B.lower" = c(KM.lower[2], COX.lower[2], NA, NA),
    "B.upper" = c(KM.upper[2], COX.upper[2], NA, NA),
    row.names = c("KM", "COX", "AFT", "AFT.STRAT")
  )
  return(tbl)
}
(true1826 <- c("A" = 1 - pweibull(1826, parameters[["A"]][[1]], parameters[["A"]][[2]]), 
               "B" = 1 - pweibull(1826, parameters[["B"]][[1]], parameters[["B"]][[2]])))
```

```{r}
predict_surv_imputed(1826, models.day)
```

## If we take the upper limit as exact time for register 2

```{r}
models.upp <- fit_imputed(expression(Surv(obs, status == "Dead")), df.obs)
```

```{r}
predict_surv_imputed(1826, models.upp)
```

## If we take the upper limit - 15 (middle of the month) as exact time for register 2

```{r}
df.obs["obs.mid"] <- ifelse(df.obs$register == "2" & df.obs$status == "Dead", 
                            df.obs$obs - 15, df.obs$obs)
models.mid <- fit_imputed(expression(Surv(obs.mid, status == "Dead")), df.obs)
```

```{r}
predict_surv_imputed(1826, models.mid)
```

## Interval censoring methods

```{r}
fit_interval <- function(Surv.obj, df, tr.cols = cols) {
  tr.names <- names(tr.cols)
  
  # 1. Non-Parametric Maximum Likelihood Estimator (NPMLE) aka Turnbull’s Estimator
  fit.NPMLE <- ic_np(eval(Surv.obj) ~ treatment, data = df)
  plot(fit.NPMLE, main = "NPMLE curves", col = tr.cols,
     xlab = "Days", ylab = "Survival")
  points(df$time, df$survival, cex = 0.1, col = "#99999966")
  legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
  
  # 2. Semi-Parametric proportional hazards model
  #fit.semi.ph <- ic_sp(eval(Surv.obj) ~ treatment, data = df, model = "ph")
  
  # 3. Semi-Parametric proportional odds model
  #fit.semi.po <- ic_sp(eval(Surv.obj) ~ treatment, data = df, model = "po")
  
  # 4. Weibull AFT/PH
  fit.aft <- survreg(eval(Surv.obj) ~ treatment, data = df, dist = "weibull")
  
  # 5. Weibull AFT/PH stratified
  fit.aft.strat <- survreg(eval(Surv.obj) ~ strata(treatment), data = df, dist = "weibull")
  
  # 6. Weibull AFT/PH icenreg
  fit.aft.ic <- ic_par(eval(Surv.obj) ~ treatment, data = df, model = "aft", 
                       dist = "weibull", weights = NULL)
  
  return(list(fit.NPMLE, NA, NA, fit.aft, fit.aft.strat, fit.aft.ic))
}
```

Fit the models

```{r}
models.int <- fit_interval(expression(Surv(lower, upper, type = "interval2")), df.obs)
```

Predict survival probabilities with 95% confidence intervals

```{r}
predict_surv_interval <- function(D, models) {
  survA <- models[[1]]$scurves$A
  survB <- models[[1]]$scurves$B
  NPMLE <- c(survA$S_curves[[1]][survA$Tbull_ints[, 2] >= D][1],
             survB$S_curves[[1]][survB$Tbull_ints[, 2] >= D][1])
  AFT <- 1 - pweibull(D, 
                      shape = 1 / models[[4]]$scale,
                      scale = c(exp(models[[4]]$coefficients[[1]]),
                                exp(sum(models[[4]]$coefficients))))
  AFT.STRAT <- 1 - pweibull(D,
                            shape = 1 / models[[5]]$scale,
                            scale = exp(models[[5]]$coefficients))
  coeffs <- exp(models[[6]]$coefficients)
  AFT.IC <- 1 - getFitEsts(models.int[[6]], 
                           newdata = data.frame(treatment = c("A", "B")), 
                           q = 1826)
  tbl <- data.frame(
    "A" = c(NPMLE[1], AFT[1], AFT.STRAT[1], AFT.IC[1]),
    "A.lower" = c(NA, NA, NA, NA),
    "A.upper" = c(NA, NA, NA, NA),
    "B" = c(NPMLE[2], AFT[2], AFT.STRAT[2], AFT.IC[2]),
    "B.lower" = c(NA, NA, NA, NA),
    "B.upper" = c(NA, NA, NA, NA),
    row.names = c("NPMLE", "AFT", "AFT.STRAT", "AFT.IC")
  )
  return(tbl)
}
(true1826 <- c("A" = 1 - pweibull(1826, parameters[["A"]][[1]], parameters[["A"]][[2]]), 
               "B" = 1 - pweibull(1826, parameters[["B"]][[1]], parameters[["B"]][[2]])))
```

```{r}
predict_surv_interval(1826, models.int)
```

## Proposed method

```{r}


```


# Comparisons



