---
title: "Full Simulation"
author: "Ugne Milasiunaite"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("./functions.R")
set.seed(311)
```

## Data

We have assumed the data comes...

```{r}
# Parameters describing the setting
N <- 5000
N.prop <- 1/2
sizes <- c("1" = N*N.prop, "2" = N - N*N.prop)
treat.prop <- list("1" = list("A" = 0.5, "B" = 0.5), 
                   "2" = list("A" = 0.5, "B" = 0.5))
lost.prop <- c("1" = 0.01, "2" = 0.01)
duration <- 365
followup <- 1826
```

Now, we create the observed data sets.

```{r}
# Proportional hazards
pars.ph <- list("A" = c(0.5, 2000), "B" = c(0.5, 2600))

# Create the data set
df.ph <- generate_df(sizes, treat.prop, pars.ph, duration, followup, lost.prop)
df.ph["obs"] <- ifelse(
  df.ph$register == 1,
  df.ph$exit,
  ifelse(df.ph$status == "Dead", df.ph$upper, df.ph$exit)
)
# Midpoint imputation (needed for later)
df.ph["mid"] <- ifelse(
  df.ph$register == "2" & df.ph$status == "Dead", 
  df.ph$obs - 15, 
  df.ph$obs
)

# Inspect the data set
df.ph[c((N1/2-20):(N1/2-19), (N1-19):(N1-20), (N1+19):(N1+20), (N-20):(N-19)),]
```

```{r}
# Non-proportional hazards
pars.nph <- list("A" = c(0.9, 2000), "B" = c(0.5, 2500))

# Create the data set
df.nph <- generate_df(sizes, treat.prop, pars.nph, duration, followup, lost.prop)
df.nph["obs"] <- ifelse(
  df.nph$register == 1,
  df.nph$exit,
  ifelse(df.nph$status == "Dead", df.nph$upper, df.nph$exit)
)
# Midpoint imputation (needed for later)
df.nph["mid"] <- ifelse(
  df.nph$register == "2" & df.nph$status == "Dead", 
  df.nph$obs - 15, 
  df.nph$obs
)

# Inspect the data set
df.nph[c((N1/2-20):(N1/2-19), (N1-19):(N1-20), (N1+19):(N1+20), (N-20):(N-19)),]
```

Now, we look at the proportional hazards assumption (compare the two data sets).

```{r}
tr.cols <- c("A" = "#009292", "B" = "#B66DFF")
tr.names <- c("A", "B")

# Fit proportional hazard data
KM.ph <- survfit(Surv(exit, status == "Dead") ~ treatment, data = df.ph, conf.type = "log-log")
COX.ph <- coxph(Surv(exit, status == "Dead") ~ treatment, data = df.ph)

# Fit non-proportional hazard data
KM.nph <- survfit(Surv(exit, status == "Dead") ~ treatment, data = df.nph, conf.type = "log-log")
COX.nph <- coxph(Surv(exit, status == "Dead") ~ treatment, data = df.nph)
```

Plot the Kaplan-Meier curves.

```{r}
par(mfcol = c(1, 2))
plot(KM.ph, main = "Kaplan-Meier curves (PH)", col = tr.cols,
     xlab = "Days", ylab = "Survival")
points(df.ph$time, df.ph$survival, cex = 0.1, col = "#99999966")
legend("bottomleft", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)

plot(KM.nph, main = "Kaplan-Meier curves (NPH)", col = tr.cols,
     xlab = "Days", ylab = "Survival")
points(df.nph$time, df.nph$survival, cex = 0.1, col = "#99999966")
legend("bottomleft", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
```

Plot the log-log plots.

```{r}
par(mfcol = c(1, 2))
plot(KM.ph, fun = "cloglog", col = tr.cols, xlab = "Days", ylab = "log(-log(Survival))", main = "PH")
legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)

plot(KM.nph, fun = "cloglog", col = tr.cols, xlab = "Days", ylab = "log(-log(Survival))", main = "NPH")
legend("bottomright", lty = c(1, 1), bty = "n", col = tr.cols, legend = tr.names)
```

Plot Schoenfeld residuals.

```{r}
par(mfcol = c(1, 2))
res.ph <- cox.zph(COX.ph, terms = TRUE)
plot(res.ph, resid = TRUE, main = "Schoenfeld residuals (PH)", cex = 0.2)

res.nph <- cox.zph(COX.nph, terms = TRUE)
plot(res.nph, resid = TRUE, main = "Schoenfeld residuals (NPH)", cex = 0.2)
```


## Model fitting


```{r}
model.names <- c("KM.DAILY", "COX.DAILY", "W.DAILY", "W.STR.DAILY",
                  "KM.IGN", "COX.IGN", "W.IGN", "W.STR.IGN",
                  "KM.IMP", "COX.IMP", "W.IMP", "W.STR.IMP",
                  "NPMLE", "SEMI.PH", "SEMI.PO", "W.IC", "W.STR.IC",
                  "SPL30", "SPL21", "W.FPM", "W.FPM1",
                  "PA.W", "PA.GAMMA", "PA.GEN")
properties <- c("A", "A.lower", "A.upper", "A.SE", 
                "B", "B.lower", "B.upper", "B.SE", 
                "AIC")
M <- c(500, 1000, 2000, 5000, 10000)
n.M <- length(M)
tables.ph <- array(NA, dim = c(n.M, 24, 9), dimnames = list(M, model.names, properties))
tables.nph <- array(NA, dim = c(n.M, 24, 9), dimnames = list(M, model.names, properties))

for(i in 1:n.M) {
  sizes <- c("1" = M[i]*N.prop, "2" = M[i] - M[i]*N.prop)
  # Proportional hazards
  df.pars.ph <- list(sizes, treat.prop, pars.ph, duration, followup, lost.prop)
  tables.ph[i,,] <- get_table(1826, df.pars.ph)
  
  # Non-proportional hazards
  df.pars.nph <- list(sizes, treat.prop, pars.nph, duration, followup, lost.prop)
  tables.nph[i,,] <- get_table(1826, df.pars.nph)
}
```


```{r}
ph1826 <- c("A" = 1 - pweibull(1826, pars.ph[["A"]][[1]], pars.ph[["A"]][[2]]), 
            "B" = 1 - pweibull(1826, pars.ph[["B"]][[1]], pars.ph[["B"]][[2]]))
nph1826 <- c("A" = 1 - pweibull(1826, pars.nph[["A"]][[1]], pars.nph[["A"]][[2]]), 
             "B" = 1 - pweibull(1826, pars.nph[["B"]][[1]], pars.nph[["B"]][[2]]))

biasA.ph <- tables.ph[,,1] - ph1826[["A"]]
biasB.ph <- tables.ph[,,5] - ph1826[["B"]]
seA.ph <- tables.ph[,,4]
seB.ph <- tables.ph[,,8]

biasA.nph <- tables.nph[,,1] - nph1826[["A"]]
biasB.nph <- tables.nph[,,5] - nph1826[["B"]]
seA.nph <- tables.nph[,,4]
seB.nph <- tables.nph[,,8]

col.M <- c("red", "darkorange", "forestgreen", "blue", "purple")
```

Now plot the bias.

```{r}
plot_bias(biasA.ph, col.M, " (A, PH)")
plot_bias(biasB.ph, col.M, " (B, PH)")
plot_bias(biasA.nph, col.M, " (A, NPH)")
plot_bias(biasB.nph, col.M, " (B, NPH)")
```

Plot AICs.

```{r}
AICs.ph <- apply(tables.ph[,,9], 1, function(v) v-min(v, na.rm = TRUE))
plot_AIC(AICs.ph, c(2:4, 6:8, 10:12), col.M, "difference from minimum")
plot_AIC(AICs.ph, 16:24, col.M, "difference from minimum")
AICs.nph <- apply(tables.nph[,,9], 1, function(v) v-min(v, na.rm = TRUE))
plot_AIC(AICs.nph, c(2:4, 6:8, 10:12), col.M, "difference from minimum")
plot_AIC(AICs.nph, 16:24, col.M, "difference from minimum")
```

Now try to estimate average bias, average AIC and SE.

```{r}
n.iter <- 20
sizes <- c("1" = N*N.prop, "2" = N - N*N.prop)
df.pars.ph <- list(sizes, treat.prop, pars.ph, duration, followup, lost.prop)
df.pars.nph <- list(sizes, treat.prop, pars.nph, duration, followup, lost.prop)

tables.ph.iter <- array(NA, dim = c(n.iter, 24, 9), dimnames = list(1:n.iter, model.names, properties))
tables.nph.iter <- array(NA, dim = c(n.iter, 24, 9), dimnames = list(1:n.iter, model.names, properties))

# Add a progress bar to help estimate how long the for loop would take
pb <- txtProgressBar(min = 0, max = n.iter, style = 3, width = 50, char = "=")
for(i in 1:n.iter) {
  
  # Proportional hazards
  #tables.ph.iter[i,,] <- get_table(1826, df.pars.ph)
  
  # Non-proportional hazards
  tables.nph.iter[i,,] <- get_table(1826, df.pars.nph)
  setTxtProgressBar(pb, i)
}
```


Now, we find the average values.

```{r}
summary.names <- c("Min.", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")
avgAIC <- apply(tables.nph.iter[,,9], 2, function(v) as.numeric(summary(v, na.rm = TRUE)))
avgBiasA <- apply(tables.nph.iter[,,1]-nph1826[["A"]], 2, function(v) round(as.numeric(summary(v, na.rm = TRUE)), 5))
avgBiasB <- apply(tables.nph.iter[,,5]-nph1826[["B"]], 2, function(v) round(as.numeric(summary(v, na.rm = TRUE)), 5))
sdA <- apply(tables.nph.iter[,,1], 2, sd, na.rm = TRUE)
sdB <- apply(tables.nph.iter[,,5], 2, sd, na.rm = TRUE)
cbind(sdA, sdB)
```

Look at average AIC, sorted by mean.

```{r}
tavgAIC <- t(as.data.frame(avgAIC[c(2:4, 6:8, 10:12, 16:24)], row.names = summary.names))
tavgAIC[order(tavgAIC[,4]),]
```

Look at bias for treatment A, sorted by mean.

```{r}
tavgBiasA <- t(as.data.frame(avgBiasA[-14], row.names = summary.names))
tavgBiasA[order(abs(tavgBiasA[,4])),]
```

```{r}
tavgBiasB <- t(as.data.frame(avgBiasB[-14], row.names = summary.names))
tavgBiasB[order(abs(tavgBiasB[,4])),]
```

