---
title: "Simulation"
author: "Ugne Milasiunaite"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we load required libraries.

```{r}
library(survival)
library(icenReg)
```

## Data

We create a data set with columns:

-   id
-   country
-   treatment
-   entry date
-   status
-   exit date (if known)
-   exit year
-   exit month

```{r}
# Helper functions

#' Get the date of the first day of the given date's month
#' 
#' @param ymd Date
#' @return Date
month_start <- function(ymd) as.Date(gsub("[0-9]{2}$", "01", ymd))


#' Get the date of the last day of the given date's month
#' 
#' @param ymd Date
#' @return Date
month_end <- function(ymd) month_start(ymd + 32) - 1


#' Stratified distribution function with no explanatory variables
#' 
#' @param v Random vector
#' @param x Treatment indicator
#' @return Vector of cumulative probabilities
pdist_strat <- function(v, x, pars) {
  lst <- sapply(names(pars), function(tr) 
    pweibull(v[x == tr], shape = pars[[tr]][[1]], scale = pars[[tr]][[2]]))
  return(do.call(c, lst))
}


#' Stratified hazard function with no explanatory variables
#' 
#' @param v Random vector
#' @param x Treatment indicator
#' @return Vector of hazards
hazard_strat <- function(v, x, pars) {
  lst <- sapply(names(pars), function(tr)
    dweibull(v[x == tr], pars[[tr]][[1]], pars[[tr]][[2]]) / (1 - pdist_strat(v[x == tr], tr, pars)))
  return(do.call(c, lst))
}


#' Generate time-to-event data from the Weibull distribution
#' 
#' @param n A named list providing the registry sizes
#' @param proportions A named list of named lists of the proportions of people receiving each treatment for each registry
#' @param pars A named list of Weibull dist. parameter values for each treatment
#' @return A data frame with four columns: survival time, probability, treatment type, registry
generate_weibull <- function(n, proportions, pars) {
  # Save names for later use
  regs <- names(proportions)
  treatments <- names(pars)

  # Calculation of sizes
  N <- sapply(regs, function(i)
    sapply(treatments, function(tr)
      round(n[[i]] * proportions[[i]][[tr]])))
  
  # Generate time-to-event data
  t.vec <- sapply(treatments, function(i) 
    rweibull(sum(N[i,]), shape = pars[[i]][[1]], scale = pars[[i]][[2]]))
  
  # Create the combined data set
  df <- do.call(rbind, lapply(treatments, function(tr)
    data.frame(
      "time" = t.vec[[tr]],
      "survival" = 1 - pweibull(t.vec[[tr]], pars[[tr]][[1]], pars[[tr]][[2]]),
      "treatment" = as.factor(tr),
      "registry" = as.factor(rep(regs, N[tr, ]))
    )))
  
  return(df)
}


#' Generate status, entry and exit dates given survival time in days
#' 
#' @param days Vector of survival time in days
#' @param start Study start date
#' @param duration Duration of follow-up from the study start in days
#' @param prob Probability of being lost-to-follow-up for each individual
#' @return A data frame with
generate_dates <- function(days, start, duration, prob) {
  # Generate entry times and 
  N <- length(days)
  lost <- rbinom(N, 1, prob)
  t0 <- floor(runif(N, 0, duration - 1))
  
  # Create the data frame
  df <- data.frame(
    "entry" = start + t0,
    "status" = as.factor(ifelse(
      lost, "Lost to follow up", ifelse(t0 + days > duration, "Alive", "Dead")
    )))
  df["exit"] <- start + ifelse(
    df$status == "Alive",
    duration,
    t0 + ifelse(!lost, days, floor(pmin(duration - t0, days)*runif(N)))
  )
  
  return(df)
}


#' Generate lower and upper limits for interval-censored data
#' 
#' @param exit Date of exit
#' @param exact Binary indicator whether the date should be left exact
#' @param lower Vector of lower limits for the lower limit
#' @param upper Vector of upper limits for the upper limit
#' @param infinite Binary indicator if the upper limit is infinite
#' @return
generate_intervals <- function(exit, exact, lower, upper, infinite) {
  L <- as.Date(ifelse(exact, exit, pmax(month_start(exit), lower)))
  U <- as.Date(ifelse(exact, exit, ifelse(infinite, Inf, pmin(month_end(exit), upper))))
  return(data.frame("lower" = L, "upper" = U))
}


#' Generate the data set
#' 
#' @param n A named list providing the registry sizes
#' @param proportions A named list of named lists of the proportions of people receiving each treatment for each registry
#' @param pars A named list of Weibull dist. parameter values for each treatment
#' @param start Study start date
#' @param duration Duration of follow-up from the study start in days
#' @param prob Probability of being lost-to-follow-up in each registry
generate_df <- function(n, proportions, pars, start, duration, prob) {
  df <- generate_weibull(n, proportions, pars)
  df_dates <- generate_dates(floor(df$time), start, duration, prob[df$registry])
  df_intervals <- generate_intervals(
    df_dates$exit,
    exact = df$registry == "1" & df_dates$status == "Dead",
    lower = df_dates$entry,
    upper = start + duration,
    infinite = df_dates$status != "Dead"
  )
  return(do.call(cbind, list(df, df_dates, df_intervals)))
}

```

We now generate the data.

```{r}
set.seed(505)

# Description of the setting
sizes <- list("1" = 1200, "2" = 900)
treat.prop <- list("1" = list("A" = 0.3, "B" = 0.2, "C" = 0.5), 
                   "2" = list("A" = 0.47, "B" = 0.36, "C" = 0.17))
parameters <- list("A" = c(0.8, 500), "B" = c(0.9, 500), "C" = c(0.7, 500))
lost.prop <- c("1" = 0.014, "2" = 0.018)
duration <- 1095
start <- as.Date("2001-01-01")

# Create the observed data set
df.obs <- generate_df(sizes, treat.prop, parameters, start, duration, lost.prop)
head(df.obs)
```

Visualize the true distributions.

```{r}
cols <- c("A" = "#009292", "B" = "#B66DFF", "C" = "#920000")
plot(df.obs$time, df.obs$survival, cex = 0.1, xlab = "Days", ylab = "Survival", 
     xlim = c(0, 2000), col = cols[df.obs$treatment])
legend("topright", lty = c(1, 1, 1), bty = "n", col = cols, legend = names(cols))
```

Plot the hazard functions.

```{r}
plot(df.obs$time, hazard_strat(df.obs$time, df.obs$treatment, parameters), cex = 0.1,
     xlim = c(0, 2000), ylim = c(0, 0.01), xlab = "Days", ylab = "Hazard",
     col = cols[df.obs$treatment])
legend("topright", lty = c(1, 1, 1), bty = "n", col = cols, legend = names(cols))
```

Add additional covariates.

```{r}
N <- nrow(df.obs)
df.obs["sex"] <- factor(rbinom(N, 1, ifelse(df.obs$time > 1000, 2/3, 1/3)), 
                        levels = 0:1, labels = c("Male", "Female"))
df.obs["smoker"] <- rbinom(N, 1, ifelse(df.obs$sex == "Male", 5/7, 1/3))
df.obs["age.cont"] <- rnorm(N, mean = ifelse(df.obs$time > 1000, 45, 60), sd = 7)
df.obs["age"] <- floor(df.obs$age)
```

## Proposed approach

```{r, warning = FALSE}
t <- df.obs[df.obs$status == "Dead", "time"]
prob <- df.obs[df.obs$status == "Dead", "survival"]
xt <- df.obs[df.obs$status == "Dead", "treatment"]

# Helper functions
fy <- function(l, u, b, x) Fz(u, b, x) - Fz(l, b, x)
logL <- function(b, l, u, x) -sum(log(fy(l, u, b, x)))

# Create coarsening intervals
l <- as.numeric(ifelse(df.obs$lower - df.obs$entry == 0, 0, df.obs$lower - df.obs$entry - 1))
u <- ifelse(df.obs$status == "Dead", df.obs$upper - df.obs$entry + 0.5, Inf)
# save covariate vector
x <- df.obs$treatment

# Try different distributions
Fz <- function(z, b, x) {
  (x == "A")*pexp(z, b[1]) + (x == "B")*pexp(z, b[2]) + (x == "C")*pexp(z, b[3])
}
mle.exp <- nlm(logL, c(1/100, 1/100, 1/100), l, u, x, hessian = TRUE)
est.exp <- 1 - Fz(t, mle.exp$estimate, xt)

Fz <- function(z, b, x) {
  (x == "A")*pweibull(z, exp(b[1]), exp(b[2])) + (x == "B")*pweibull(z, exp(b[3]), exp(b[4])) + (x == "C")*pweibull(z, exp(b[5]), exp(b[6]))
  }
mle.weibull <- nlm(logL, c(-0.2, 6, -0.2, 6, -0.2, 6), l, u, x, hessian = TRUE)
est.weibull <- 1 - Fz(t, mle.weibull$estimate, xt)

Fz <- function(z, b, x) {
  (x == "A")*pgamma(z, b[1], b[2]) + (x == "B")*pgamma(z, b[3], b[4]) + (x == "C")*pgamma(z, b[5], b[6])
}
mle.gamma <- nlm(logL, c(1, 1/100, 1, 1/100, 1, 1/100), l, u, x, hessian = TRUE)
est.gamma <- 1 - Fz(t, mle.gamma$estimate, xt)

# Plot and compare the distributions
plot(t, prob, cex = 0.1, xlab = "Days", ylab = "Survival probability", 
     main = "Fitted survival curves")
points(t, est.exp, col = "#0000FF66", cex = 0.1)
points(t, est.weibull, col = "#72BF4066", cex = 0.1)
points(t, est.gamma, col = "#FF000066", cex = 0.1)
legend("topright", lty = rep(1, 4), bty = "n",
       col = c("black", "#0000FF99", "#72BF4099", "#FF000099"), 
       legend = c("True distribution", "Exponential", "Weibull", "Gamma"))

```

Proposed approach, generalized

```{r, warning = FALSE}
# define the pre-transform
s <- function(z) qlogis(pweibull(z, 0.8, 500))
# apply the pre-transform to the coarsening limits
sl <- s(l)
su <- s(u)
# define the modelling distribution and polynomial approximation
Fz <- function(z, b, x) plogis(g(z, b, x))
g <- function(z, b, x) {
  (b[1] + b[2]*(x == "B") + b[3]*(x == "C")) + z*(b[4] + b[5]*(x == "B") + b[6]*(x == "C"))
}
fy <- function(l, u, b, x) Fz(u, b, x) - Fz(l, b, x)
logL <- function(b, l, u, x) -sum(log(fy(l, u, b, x)))
#
pdist_ge <- function(z, b, x) plogis(g(s(z), b, x))
# estimate the parameters
pars <- nlm(logL, rep(1, 6), u=su, l=sl, x=df.obs$treatment)
est.ge <- 1 - Fz(s(t), pars$estimate, xt)
# plot the estimated survival distributions
plot(t, prob, col="black", cex = 0.1, xlab = "Days", ylab = "Survival")
points(t, est.weibull, col = "#72BF4066", cex = 0.1)
points(t, est.ge, col = "navy", cex = 0.1)
```

Proposed approach, accelerated failure time

```{r, warning = FALSE}
# Need to look at this later
fy <- function(l, u, b, x) Fz(u, b, x) - Fz(l, b, x)
logL <- function(b, l, u, x) -sum(log(fy(l, u, b, x)))
x <- df.obs$treatment
Fz <- function(z, b, x) {
  (x == "A")*pweibull(z*exp(-b[3]), exp(b[1] + b[3]*b[2]), exp(b[2])) + (x == "B")*pweibull(z*exp(-b[6]), exp(b[4] + b[6]*b[5]), exp(b[5])) + (x == "C")*pweibull(z*exp(-b[9]), exp(b[7] + b[9]*b[8]), exp(b[8]))
  }
pars.aft <- nlm(logL, c(-0.69, 6, 0, -0.69, 6, 0, -0.69, 6, 0), l, u, x, hessian = TRUE)
est.aft <- 1 - Fz(t, pars.aft$estimate, xt)
plot(t, prob, col = "black", cex = 0.1, xlab = "Days", ylab = "Survival")
points(t, est.weibull, col = "#72BF4066", cex = 0.1)
points(t, est.ge, col = "navy", cex = 0.1)
points(t, est.aft, col = "maroon", cex = 0.1)
```

Some numerical comparisons

```{r}
mean((prob - est.gamma)^2)
mean((prob - est.weibull)^2)
mean((prob - est.ge)^2)
mean((prob - est.aft)^2)
print("-------")
mean(abs(prob - est.gamma))
mean(abs(prob - est.weibull))
mean(abs(prob - est.ge))
mean(abs(prob - est.aft))
```

## Imputation

```{r}
df.obs["mid"] <- as.Date(ifelse(
  df.obs$upper == df.obs$lower | df.obs$status != "Dead",
  df.obs$lower,
  df.obs$lower + floor((df.obs$upper - df.obs$lower)/2)))
df.obs["days.mid"] <- as.numeric(df.obs$mid - df.obs$entry)

# Fit the Kaplan-Meier survival curve
KM <- survfit(Surv(days.mid, status == "Dead") ~ treatment, data = df.obs, 
              conf.type = "log-log")
v <- KM$time  # unique time points for which we can estimate the probabilities

# Compare the Kaplan-Meier curve with the true distribution
plot(KM, col = "red", main = "Kaplan-Meier survival curve", 
     xlab = "Days", ylab = "Survival")
points(t, prob, cex = 0.1, col = "black")
```

Compare with the previous approach.

```{r}
plot(KM, main = "Kaplan-Meier vs proposed approach", xlab = "Days", ylab = "Survival")
points(t, est.weibull, col = "#72BF4066", cex = 0.1)
points(t, est.gamma, col = "#FF000066", cex = 0.1)
legend("topright", lty = rep(1, 3), bty = "n",
       col = c("black", "#72BF4099", "#FF000099"), 
       legend = c("Kaplan-Meier curve", "Weibull", "Gamma"))
```

Compare with a parametric approach.

```{r}
ta <- t[xt == "A"]
tb <- t[xt == "B"]
tc <- t[xt == "C"]
fit.weibull <- survreg(Surv(days.mid, status == "Dead") ~ strata(treatment), 
                       data = df.obs[df.obs$days.mid != 0,], dist = "weibull")
fit.weibull.probs <- c(
  pweibull(ta, shape = 1/exp(fit.weibull$icoef[[2]]), 
           scale = exp(fit.weibull$icoef[[1]])), 
  pweibull(tb, shape = 1/exp(fit.weibull$icoef[[3]] + fit.weibull$icoef[[2]]), 
           scale = exp(fit.weibull$icoef[[1]])),
  pweibull(tc, shape = 1/exp(fit.weibull$icoef[[4]] + fit.weibull$icoef[[2]]), 
           scale = exp(fit.weibull$icoef[[1]]))
)

plot(c(ta, tb, tc), 1 - fit.weibull.probs, xlab = "Days", ylab = "Survival", 
     col = "blue", main = "Stratified Weibull parametric model", cex = 0.1)
#points(t, est.weibull, col = "#72BF4066", cex = 0.1)
#points(t, est.gamma, col = "#FF000066", cex = 0.1)
points(t, prob, col = "black", cex = 0.1)
legend("topright", lty = rep(1, 4), bty = "n",
       col = c("black", "blue", "#72BF4099", "#FF000099"),
       legend = c("True distribution", "Parametric Weibull", "Weibull", "Gamma"))
```

```{r}
# Mean squared errors
mean((prob - est.gamma)^2)
mean((prob - est.weibull)^2)
mean((prob - est.exp)^2)
mean((c(prob[xt == "A"], prob[xt == "B"], prob[xt == "C"]) - 1 + fit.weibull.probs)^2)
mean((1 - pdist_strat(KM$time, rep(c("A", "B", "C"), KM$strata), parameters) - KM$surv)^2)
```

```{r}
# Absolute errors
mean(abs(prob - est.gamma))
mean(abs(prob - est.weibull))
mean(abs(prob - est.exp))
mean(abs(c(prob[xt == "A"], prob[xt == "B"], prob[xt == "C"]) - 1 + fit.weibull.probs))
mean(abs(1 - pdist_strat(KM$time, rep(c("A", "B", "C"), KM$strata), parameters) - KM$surv))
```

## Check for proportional hazards assumption

```{r}
# Plot hazards using muhaz
library(muhaz)
df.obs.A <- df.obs[df.obs$treatment == "A",]
df.obs.B <- df.obs[df.obs$treatment == "B",]
df.obs.C <- df.obs[df.obs$treatment == "C",]
mhA.mid <- muhaz(df.obs.A$days.mid, df.obs.A$status == "Dead")
mhB.mid <- muhaz(df.obs.B$days.mid, df.obs.B$status == "Dead")
mhC.mid <- muhaz(df.obs.C$days.mid, df.obs.C$status == "Dead")
mhA.l <- muhaz(df.obs.A$lower - df.obs.A$entry, df.obs.A$status == "Dead")
mhB.l <- muhaz(df.obs.B$lower - df.obs.B$entry, df.obs.B$status == "Dead")
mhC.l <- muhaz(df.obs.C$lower - df.obs.C$entry, df.obs.C$status == "Dead")
mhA.u <- muhaz(month_end(df.obs.A$lower) - df.obs.A$entry, df.obs.A$status == "Dead")
mhB.u <- muhaz(month_end(df.obs.B$lower) - df.obs.B$entry, df.obs.B$status == "Dead")
mhC.u <- muhaz(month_end(df.obs.C$lower) - df.obs.C$entry, df.obs.C$status == "Dead")
# Plot the estimated curves
par(mfrow = c(3, 1), pin = c(5, 0.9), mar = c(1, 2, 1, 1), mai = c(0.25, 0.5, 0.15, 0.1))
plot(mhA.mid, ylim=c(0,0.007), col = cols["A"], main = "Midpoint imputation", xaxt = "n")
lines(mhB.mid, col = cols["B"])
lines(mhC.mid, col = cols["C"])
legend("topleft", lty = c(1, 1, 1), bty = "n", col = cols, 
       legend = c("A", "B", "C"))
plot(mhA.l, ylim=c(0,0.007), col = cols["A"], main = "Month start imputation", xaxt = "n")
lines(mhB.l, col = cols["B"])
lines(mhC.l, col = cols["C"])
legend("topleft", lty = c(1, 1, 1), bty = "n", col = cols, 
       legend = c("A", "B", "C"))
plot(mhA.u, ylim=c(0,0.007), col = cols["A"], main = "Month end imputation")
lines(mhB.u, col = cols["B"])
lines(mhC.u, col = cols["C"])
legend("topleft", lty = c(1, 1, 1), bty = "n", col = cols, 
       legend = c("A", "B", "C"))
```


```{r}
fit.cox <- coxph(Surv(days.mid, status == "Dead") ~ treatment, data = df.obs)
res <- cox.zph(fit.cox, terms = FALSE)
print(res)
```

```{r}
plot(res, resid = TRUE)
```

The test indicates that the assumption of proportional hazards is violated. By plotting the Schoenfeld residuals we see that the hazards seem proportional until around day 410, at which point the points start to go up and diverge from a horizontal line.

## Methods for interval-censored data

```{r}
df.obs["days.lower"] <- as.numeric(df.obs$lower - df.obs$entry)
df.obs["days.upper"] <- as.numeric(df.obs$upper - df.obs$entry)
fit.np <- ic_np(cbind(days.lower, days.upper) ~ treatment, data = df.obs)
plot(fit.np)
points(t, prob, col = "black", cex = 0.1)
```

```{r}
df.obs["days.lower"] <- ifelse(df.obs$days.lower == 0 & df.obs$days.upper == 0, 
                               0.5, df.obs$days.lower)
df.obs["days.upper"] <- ifelse(df.obs$days.upper == 0, 0.5, df.obs$days.upper)
fit.par <- ic_par(cbind(days.lower, days.upper) ~ treatment, data = df.obs, 
                  model = "aft", dist = "weibull", weights = NULL)
plot(fit.par, col = "red", xlim = c(0, 1000))
plot(ta, 1 - pweibull(ta, shape = exp(fit.par$coefficients[[1]]), scale = exp(fit.par$coefficients[[2]] - fit.par$coefficients[[3]])), col = cols["A"], cex = 0.1, xlab = "Days", ylab = "Survival")
points(tb, 1 - pweibull(tb, shape = exp(fit.par$coefficients[[1]]), scale = exp(fit.par$coefficients[[2]])), col = cols["B"], cex = 0.1)
points(tc, 1 - pweibull(tc, shape = exp(fit.par$coefficients[[1]]), scale = exp(fit.par$coefficients[[2]] - fit.par$coefficients[[4]])), col = cols["C"], cex = 0.1)
#points(t, prob, col = "black", cex = 0.1)
```

Check the fit

```{r}
mean((c(fit.np$scurves$A$S_curves$baseline - 1 + pdist_strat(fit.np$scurves$A$Tbull_ints[, 1], "A", parameters), fit.np$scurves$B$S_curves$baseline - 1 + pdist_strat(fit.np$scurves$B$Tbull_ints[, 1], "B", parameters), fit.np$scurves$C$S_curves$baseline - 1 + pdist_strat(fit.np$scurves$C$Tbull_ints[, 1], "C", parameters)))^2)
mean(abs(c(fit.np$scurves$A$S_curves$baseline - 1 + pdist_strat(fit.np$scurves$A$Tbull_ints[, 1], "A", parameters), fit.np$scurves$B$S_curves$baseline - 1 + pdist_strat(fit.np$scurves$B$Tbull_ints[, 1], "B", parameters), fit.np$scurves$C$S_curves$baseline - 1 + pdist_strat(fit.np$scurves$C$Tbull_ints[, 1], "C", parameters))))
```

```{r}
diag_baseline(fit.par)
```


## Make a table of survival probabilities

```{r}
# "A"
days <- as.numeric(month_end(start + 1 + c(0, 1, 2, 3, 4, 5, 11, 17, 23, 35, 47, 59)*31) - start)
tbl.survA <- data.frame("months" = c(1, 2, 3, 4, 5, 6, 12, 18, 24, 36, 48, 60),
                       "days" = days)
tbl.survA["true"] <- 1 - pdist_strat(days, "A", parameters)
tbl.survA["weibull"] <- 1 - pweibull(days, exp(mle.weibull$estimate[1]), exp(mle.weibull$estimate[2])) - tbl.survA$true
tbl.survA["gamma"] <- 1 - pgamma(days, mle.gamma$estimate[1], mle.gamma$estimate[2]) - tbl.survA$true
tbl.survA["generalised"] <- 1 - pdist_ge(days, pars$estimate, "A") - tbl.survA$true
tbl.survA["weibull.aft"] <- 1 - Fz(days, pars.aft$estimate, "A") - tbl.survA$true
tbl.survA["imp.weibull"] <- 1 - pweibull(days, shape = 1/exp(fit.weibull$icoef[[2]]), 
           scale = exp(fit.weibull$icoef[[1]])) - tbl.survA$true
tbl.survA["int.weibull"] <- 1 - pweibull(days, shape = exp(fit.par$coefficients[[1]]), scale = exp(fit.par$coefficients[[2]])) - tbl.survA$true
head(tbl.survA)
```


```{r}
# "B"
tbl.survB <- data.frame("months" = c(1, 2, 3, 4, 5, 6, 12, 18, 24, 36, 48, 60),
                       "days" = days)
tbl.survB["true"] <- 1 - pdist_strat(days, "B", parameters)
tbl.survB["weibull"] <- 1 - pweibull(days, exp(mle.weibull$estimate[3]), exp(mle.weibull$estimate[4])) - tbl.survB$true
tbl.survB["gamma"] <- 1 - pgamma(days, mle.gamma$estimate[3], mle.gamma$estimate[4]) - tbl.survB$true
tbl.survB["generalised"] <- 1 - pdist_ge(days, pars$estimate, "B") - tbl.survB$true
tbl.survB["weibull.aft"] <- 1 - Fz(days, pars.aft$estimate, "B") - tbl.survB$true
tbl.survB["imp.weibull"] <- 1 - pweibull(days, shape = 1/exp(fit.weibull$icoef[[3]]), 
           scale = exp(fit.weibull$icoef[[1]])) - tbl.survB$true
tbl.survB["int.weibull"] <- 1 - pweibull(days, shape = exp(fit.par$coefficients[[1]]), scale = exp(fit.par$coefficients[[2]] + fit.par$coefficients[[3]]/(exp(fit.par$coefficients[[1]])-1))) - tbl.survB$true
head(tbl.survB)
```

```{r}
# "C"
tbl.survC <- data.frame("months" = c(1, 2, 3, 4, 5, 6, 12, 18, 24, 36, 48, 60),
                       "days" = days)
tbl.survC["true"] <- 1 - pdist_strat(days, "C", parameters)
tbl.survC["weibull"] <- 1 - pweibull(days, exp(mle.weibull$estimate[5]), exp(mle.weibull$estimate[6])) - tbl.survC$true
tbl.survC["gamma"] <- 1 - pgamma(days, mle.gamma$estimate[5], mle.gamma$estimate[6]) - tbl.survC$true
tbl.survB["generalised"] <- 1 - pdist_ge(days, pars$estimate, "C") - tbl.survC$true
tbl.survC["weibull.aft"] <- 1 - Fz(days, pars.aft$estimate, "C") - tbl.survC$true
tbl.survC["imp.weibull"] <- 1 - pweibull(days, shape = 1/exp(fit.weibull$icoef[[4]]), 
           scale = exp(fit.weibull$icoef[[1]])) - tbl.survC$true
tbl.survC["int.weibull"] <- 1 - pweibull(days, shape = exp(fit.par$coefficients[[1]]), scale = exp(fit.par$coefficients[[2]] + fit.par$coefficients[[4]]/(exp(fit.par$coefficients[[1]])-1))) - tbl.survC$true
head(tbl.survC)
```

# Adding covariates to the proposed approach

Try for treatment (startified) and adjusted for sex and smoking

```{r, warning = FALSE}
xt.tss <- as.factor(paste0(df.obs[df.obs$status == "Dead", "treatment"], df.obs[df.obs$status == "Dead", "sex"], df.obs[df.obs$status == "Dead", "smoker"]))

# Helper functions
fy <- function(l, u, b, x) Fz(u, b, x) - Fz(l, b, x)
logL <- function(b, l, u, x) -sum(log(fy(l, u, b, x)))

# Create coarsening intervals
l <- as.numeric(ifelse(df.obs$lower - df.obs$entry == 0, 0, 
                       df.obs$lower - df.obs$entry - 0.5))
u <- ifelse(df.obs$status == "Dead", df.obs$upper - df.obs$entry + 0.5, Inf)
# save covariate vector
x.tss <- as.factor(paste0(df.obs$treatment, df.obs$sex, df.obs$smoker))
names.tss <- 1:length(levels(x.tss))
names(names.tss) <- levels(x.tss)

# Try different distributions
Fz <- function(z, b, x) {
  pweibull(z, exp(b[names.tss[x]*2-1]), exp(b[names.tss[x]*2]))
}
mle.tss.w <- nlm(logL, rep(c(-0.69, 6), 12), l=l, u=u, x=x.tss, hessian = TRUE)
est.tss.w <- 1 - Fz(t, mle.tss.w$estimate, xt.tss)
estA.tss.w <- 1 - (Fz(t, mle.tss.w$estimate, "AFemale0") + Fz(t, mle.tss.w$estimate, "AFemale1") + Fz(t, mle.tss.w$estimate, "AMale0") + Fz(t, mle.tss.w$estimate, "AMale1"))/4
estB.tss.w <- 1 - (Fz(t, mle.tss.w$estimate, "BFemale0") + Fz(t, mle.tss.w$estimate, "BFemale1") + Fz(t, mle.tss.w$estimate, "BMale0") + Fz(t, mle.tss.w$estimate, "BMale1"))/4
estC.tss.w <- 1 - (Fz(t, mle.tss.w$estimate, "CFemale0") + Fz(t, mle.tss.w$estimate, "CFemale1") + Fz(t, mle.tss.w$estimate, "CMale0") + Fz(t, mle.tss.w$estimate, "CMale1"))/4
mu.est.tss.w <- (estA.tss.w + estB.tss.w + estC.tss.w)/3

# Plot and compare the distributions
plot(t, prob, cex = 0.1, xlab = "Days", ylab = "Survival probability", 
     main = "Adjusted survival curves")
points(t, estA.tss.w, col = cols["A"], cex = 0.1)
points(t, estB.tss.w, col = cols["B"], cex = 0.1)
points(t, estC.tss.w, col = cols["C"], cex = 0.1)
points(t, est.weibull, col = "#72BF4066", cex = 0.1)
legend("topright", lty = rep(1, 5), bty = "n",
       col = c("black", "#72BF40FF", cols), 
       legend = c("Unadjusted true distributions", "Unadjusted Weibull", "A, adjusted", "B, adjusted", "C, adjusted"))


```

Adjusted parametric models

```{r}
fit.par2 <- ic_par(cbind(days.lower, days.upper) ~ treatment + sex + smoker, 
                   data = df.obs, model = "aft", dist = "weibull", weights = NULL)
plot(fit.par2, col = "red", xlim = c(0, 1000))
points(t, mu.est.tss.w, col = "navy", cex = 0.1)
```



